Optimised Cost Considering Huffman Code for Biological Data Compression

### 2015-07-22 International Journal of Information and Communication Technology/Our Manuscript/manuscript.tex
new

\documentclass[twocolumn,12pt,a4paper]{article}
\usepackage[top=1.1in, bottom=1in, left=0.85in, right=0.85in]{geometry}

\usepackage{graphicx}
\usepackage{mathptmx}
\usepackage{abstract}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\usepackage{multirow}
\usepackage[round]{natbib}
\usepackage{moresize}
\usepackage{setspace}
%\singlespacing
\setlength\columnsep{20pt}
\setlength{\parindent}{0pt}
%\bibliographystyle{plainnat}
%\renewcommand{\baselinestretch}{1.5} 
\begin{document}

\title{\huge\textbf{Optimised Cost Considering Huffman \\Code for Biological Data Compression}}

%\author{
%~~Youcef Gheraibia$^{1,2}$, Sohag Kabir$^{1}$, Mohammad Kaykobad$^{3}$,~~~~~~~~~~~~~~~~~\\Abdelouahab Moussaoui$^{4}$, and Smaine Mazouzi$^{2}$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\
%\small $^{1}$Department of Computer Science, University of Hull, Hull, UK~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
%\small $^{2}$Department of Computer Science, Badji Mokhtar-Annaba University, Annaba, Algeria~~~\\
%\small $^{3}$Department of Computer Science and Engineering, Bangladesh University  of~~~~~~~~~~~~~~~~~ \\ \small Engineering  and Technology, Dhaka, Bangladesh~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
%\small $^{4}$LRIA laboratory, Department of Computer Science, University of Setif, Algeria~~~~~~~~~~~~}
\date{}
\maketitle

%\begin{abstract}
%The abstract text goes here.
%\end{abstract}
\noindent
 {\scriptsize  Classical Huffman code has widely been used to compress biological datasets. Though a considerable reduction of size of data is obtained by classical Huffman code, yet, more efficient encoding is possible by treating binary bits differently considering requirement of transmission time, energy consumption, and likewise. A number of techniques already modified Huffman code algorithm to  obtain optimal prefix-codes for unequal letter costs in order to reduce overall transmission cost (time). In this paper, we propose a new approach to improve compression performance of one such extension, cost considering approach (CCA), by applying genetic algorithm for optimal allocation of the codewords to the symbols. The approach starts it operation by generating a set of codewords for different input symbols based on their frequencies by considering cost (length) of 0 and 1 as $\alpha$ and $\beta$,~$\alpha < \beta$. This step ensures optimal cost encoding, however, compression performance is not better than classical Huffman code. The idea of the proposed approach is to sacrifice some cost to minimise the total number of bits, hence, the genetic algorithm works by giving penalty on the cost. The performance of the approach is evaluated by applying it to compress some standard biological datasets. The experiments yield that the proposed approach improves the compression performance of the CCA considerably without increasing the cost significantly.\\ \\
\noindent
\textit{Keywords:}
Data Compression, Huffman Code, Genetic Algorithm, Cost Considering Approach, Data Communication.}

\normalsize
\section{Introduction}
\label{sec1}
In the recent years, application of battery-powered portable devices, e.g. laptop computers and mobile phones has increased significantly. Proper representation of digital data and their transmission efficiency has become a primary concern for digital community because it affects the performance, reliability, and the cost of computation in both portable and non-portable devices. CMOS technologies were developed in order to reduce the power consumption both in data processing and transmission. In order to increase transmission speed and reduce transmission cost, parallel data transmission methods are widely used. However, parallel transmission is limited to short distance communications, e.g. locally connected devices, internal buses. Ruling out the possible availability of parallel transmission links over long distance, we are left with its serial alternative only. If we attempt to transfer big files, e.g. DNA sequences, over a serial transmission link then it would take a significant amount of time. However, we cannot overlook this problem because at present parallel processing is widely used to increase throughput and in parallel processing architecture, processing units are usually distributed in different physical locations and task sharing is a must in such architecture.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.5}
\caption{Example of binary logic level}
\label{table1}
\centering
\begin{tabular}{| c | c | c |}
\hline
 \bfseries Technology  & 0 & 1\\
\hline
\bfseries CMOS & $0~V$ to $\frac{V_{DD}}{2}$&$\frac{V_{DD}}{2}$ to $V_{DD}$\\
\hline
\bfseries TTL & $0~V$ to $0.8~V$&$2~V$ to $V_{CC}$ \\
%\hline
\hline
\end{tabular}
\end{table}

@@

A recent deluge of interest in the development of new tools for biological data processing requires efficient methods for data compression. The main objective of data compression methods is minimising the number of bits in the data representation.\\ 
\citet{bra09} proposed a new general data structure and data encoding approach for the efficient storage of genomic data. This method encodes only the differences between a genome sequence and a reference sequence. For encoding, the method uses different fixed-length encoding schemes such as Golomb \citep{Golomb96}, Elias codes \citep{Elias75} and variable codes such as Huffman codes. There are other methods based on the same idea of encoding only the difference between reference sequence and the target one. One such approach~\citep{chr09} uses Huffman code for encoding difference between sequences to sent it as an email attachment. The main limitation of this method is that it must send the reference sequence for at least once for each species, and usually this sequence is too big to be sent as an email attachment.
\citet{wan11} proposed a new scheme for referential compression of genomes based on the chromosome level. The algorithm searches for longest common subsequence between two sequences and then the differences between them are encoded using Huffman code.
All previous studies focus only on the differences and the relation between continuation of the sequence, and use existing encoding approaches to encode biological datasets without considering possible improvement of the encoding schemes.

%\subsection{Rationale of Unequal Bit Cost Considering Encoding Approaches}

\subsection{Huffman Codes}
In computer science and information theory, Huffman code is an entropy encoding algorithm used for lossless data compression. It takes into account the probabilities at which different symbols are likely to occur and results into fewer data bits on the average. 
%then assigns variable-length codes to different distinct symbols, and thus minimises the number of bits required to encode a message. 
For any given set of symbols and associated occurrence probabilities, there is an optimal encoding rule that minimises the number of bits needed to represent the source. Encoding symbols in predefined fixed length code, does not attain an optimum performance, because every character consumes equal number of bits irrespective to their degree of contribution to the whole data. Huffman code tackles this by generating variable length codes, given a probability usage frequency for a set of symbols. It generates prefix-code to facilitate unambiguous retrieval of information. A scheme of prefix code assigns codes to letters in $\Sigma$ to form codeword $w_i$ such that none of them is a prefix to another. For example, the codes $\left\{ 1,01,001,0001\right\}$ and $\left\{ 000,001,011,111\right\}$ are prefix-free, whereas the code $\left\{ 1,01,100\right\}$ is not, because 1 is a prefix in 100.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.3}
\caption{Example of application of Huffman Code to compress ASCII characters}
\label{table2}
\centering
\begin{tabular}{ | c | c | c | c |}
\hline
 Symbols & Frequency&\begin{tabular}{@{}c@{}}ASCII\\Code\end{tabular}   &\begin{tabular}{@{}c@{}}Huffman\\ Code\end{tabular} \\
\hline
\bfseries A & 50&01000001&00\\
\hline
\bfseries B & 35&01000010&101\\
\hline
\bfseries C & 42&01000011&110\\
\hline
\bfseries D & 22&01000100&1001\\
\hline
\bfseries E & 65&01000101&01\\
\hline
\bfseries F & 25&01000110&1111\\
\hline
\bfseries G & 9&01000111&1000\\
\hline
\bfseries H & 23&01001000&1110\\
\hline
\end{tabular}
\end{table}

@@

\section{The Proposed Approach}
\label{sec3}
%\subsection{}

A genome is a stretch of DNA that encodes  a polypeptide (protein) which is a set of amino acids bound together in a specific order. Each genomic sequence consists of nucleotides bound together, which are interpreted by the cellular machinery in groups of three, called triplets~\citep{Harvey00}. This is the main reason to divide the whole sequence in a set of triplet and give a code to each triplet. As each DNA sequence contains a combination of four nucleobases\textemdash guanine (G), adenine (A), thymine (T), and cytosine (C), therefore it is possible to have $4^3=64$ triplets.  The first step in the optimised cost considering algorithm is cutting the genome sequence in triplets, then compute the frequency of each triplet in the whole sequence. This table of frequencies are used by the cost considering Huffman code to generate minimal cost code for each triplet (frequency). Finally, these codes with frequencies are used by the optimised cost considering algorithm to generate the optimal allocation with a given  penalty on cost. The whole process is shown in Fig.1.%\ref{Fig1}.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.4]{Images/Drawing3_1.jpg}
\caption{The proposed scheme}
\end{center}
\label{Fig1}
\end{figure}

@@

\begin{algorithm}[!btph]
\caption{Cost Considering Algorithm (CCA)}
\label{alg1}

@@

\subsection{Optimisation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each triplet can be   modelled as an assignment problems, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $F=\left\{F_{1},F_{2}...F_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$, the objective is to assign to each frequency a code in order to minimise the total number of bits, while respecting the initially assigned total cost $S_{t}$ with a given penalty $\lambda\rightarrow [0,1]$. This penalty coefficient represents the allowed amount of cost that can be sacrificed to optimise the total number of bits.    
\\
The Objective Function is to:

@@

\subsubsection{Basic Genetic Algorithm}
Genetic Algorithm (GA) is a bio-inspired meta-heuristic algorithm developed by \citep{gen1}. GA is a stochastic optimisation algorithm imitate the natural evolution process of genomes. GA started by generating an initial population of random feasible solutions. The optimisation process of GA takes the initial population and generates a new population based on it.  The process can be described as follow:\\ 
Firstly, select two or more solutions from the current population by using one of the well-known selection techniques \citep{bli95}. These selected solutions will be considered as  parents. The second operation of the genetic algorithm is the crossover, which takes these parents as inputs to generate other new solutions considered as children. The third operation of the genetic algorithm is the mutation which ensures a good diversification in the search process. The new solutions can be mutated according to a given mutation probability. The mutation operation changes the value of one or more positions in the solution. The quality of each solution is verified by the fitness function which controls the evolution of the GA population by deletion of the worst solutions and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is reached which can be the number of generation or if the population is stabilised.

\subsubsection{Optimised cost considering algorithm}
The main objective of the optimised cost considering algorithm (OCCA)is to apply GA to find an optimal allocation of the codewords to each triplets to reduce the total number of bits. In GA, the encoding of the solution into a chromosome is one of the most important issues in obtaining good optimisation results. The OCCA uses two fixed length arrays of size ``64" which is the number of combination for all nucleotides. The first array contains the frequencies of each triplets and the second array contains the cost of the codewords assigned to each triplets. Our genetic algorithm utilises the two arrays and uses the index of each entries in the allocation process.

@@

The evolution of the population is the key of the genetic optimisation algorithm. During each generation the process start with the selection of a part (set of solutions) of the population to bread a new generation. In the literature many selection methods have been proposed to guide the population evolution \citep{bli95}. The different methods for selecting a part (set of solutions) of the population vary from a random selection method to heuristic based selection method. We have chosen to select randomly the part of the population to be processed as the heuristic methods are very time-consuming (step 3).
After that the operations of genetic algorithm is applied on the initial population to generate a new generation of the population (see Fig.2). Firstly the crossover operation is applied to these two selected solutions (considered as parents) to generate two new solutions (considered as children)(step 4). In the literature many crossover techniques have been used in genetic algorithm \citep{osa14}, such as one-point crossover which divides the chromosomes into two fragments and recombines the second fragments with the other chromosome's second fragment. The two-point crossover divides the chromosomes into three fragments and recombines the middle fragment with  the middle fragment of the other chromosome. There are many other crossover techniques to allow a good convergence of the algorithm. In our case, we have used the two point crossover with two parameters, the first parameter $p_{c}$ is a random value in [0,63], which represents the first cut point and the second parameter $p_{m}$ is also a random value in [0,63], which represents the number of positions to be crossed. These two random parameters are to ensure a good diversification on the whole search space (step 4).

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.28]{Images/Drawing2.jpg}
\caption{Operations of genetic algorithm}
\end{center}
\label{Fig3}
\end{figure}

\begin{figure}[!thpb]
\begin{center}
\includegraphics[scale=0.3]{Images/Drawing_1.jpg}
\caption{Population update for genetic algorithm}
\end{center}
\label{Fig2}
\end{figure}

@@

After the crossover operation, the newly generated children may contain conflict, for example, a single code is allocated to two different frequencies in the solution. To overcome such a conflict, a regulation operation is performed to refine the solution to ensure the correctness of the solution (see Fig.2).
 Secondly these two new solutions are mutated according to a predefined probability $\gamma$, the best value of the mutation rate is very problem specific (step 5). In our case, the value of $\gamma$ is fixed to $0.2$ to explorer a few positions in the solution. The mutation operator is used to maintain genetic diversity from one generation of the population of genetic algorithm to the next. In our case, we have used the mutation as a random swap mutation operator (see Fig.3).  Each newly generated solution must satisfy the cost constraint. The next step is to add these two new solutions (children) to the current population (step 7) (see Fig. 3). Finally, the new population are ranked by fitness (step 8), and the worst solutions are deleted until the initial size of the population is obtained  (step 9). The whole processes are repeated until the maximum number of operation is performed (step 10).

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.26]{Images/drawing4_1.jpg}
\caption{Convergence of OCCA for \textit{Genome 2}}
\end{center}
\label{Fig5}
\end{figure}

\begin{table*}[!thpb]
\renewcommand{\arraystretch}{1.3}
\small
\label{table3}
\caption{Datasets description}
%\begin{center}
\centering
\begin{tabular}{| c | c | c | c |}
\hline
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MB)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &  6.66 & CP009496  \\
\hline

Genome 2& Amycolatopsis benzoatilytica & 8.30 & NZ\_KB912942 \\
\hline

@@

\hline

Genome 4 & \begin{tabular}{@{}c@{}}Streptomyces bottropensis\\ATCC 25435\end{tabular}& 8.54 &
NZ\_KB911581\\
\hline

Genome 5 & \begin{tabular}{@{}c@{}}Mycobacterium smegmatis\\str. MC2 155\end{tabular}& 6.66 &
CP009494            \\
\hline

Genome 6 & \begin{tabular}{@{}c@{}}Mycobacterium smegmatis\\MKD8\end{tabular}& 6.76 &
NZ\_KI421511            \\
\hline

Genome 7& Bradyrhizobium WSM471&  7.42 &NZ\_CM001442\\\hline

Genome 8 & \begin{tabular}{@{}c@{}}Amycolatopsis thermoflava\\N1165\end{tabular}& 8.27 &
NZ\_CM001442 \\
\hline

Genome 9 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\Bt407\end{tabular}& 5.74 &
NZ\_CM000747 \\
\hline

Genome 10 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar thuringiensis\end{tabular}& 6.03 & NZ\_CM000748 \\
\hline

Genome 11 & \begin{tabular}{@{}c@{}}Pseudomonas aeruginosa\\9BR\end{tabular}& 6.48 & NZ\_AFXI01000001 \\
\hline

Genome 12 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar berliner ATCC\end{tabular}& 5.97 & NZ\_CM000753 \\
\hline

Genome 13 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar pakistani\end{tabular}& 5.75 & NZ\_CM000750 \\
\hline

Genome 14 & \begin{tabular}{@{}c@{}}Pseudomonas aeruginosa\\LES400\end{tabular}& 6.28 & CP006982 \\
\hline

Genome 15 & \begin{tabular}{@{}c@{}}Mus musculus\\chromosome 1\end{tabular}& 25.58 & GL456087 \\
\hline

Genome 16 & \begin{tabular}{@{}c@{}}Danio rerio\\chromosome 1\end{tabular}& 56.14 & CM002885 \\
\hline

Genome 17 & \begin{tabular}{@{}c@{}}Homo sapiens\\chromosome 18\end{tabular}& 76.64 & CM000680  \\
\hline

Genome 18 & \begin{tabular}{@{}c@{}}Homo sapiens\\chromosome 22\end{tabular}& 99.94 & CM000684    \\
\hline

%\bottomrule
\end{tabular}
%\end{center}
\end{table*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%
\documentclass{singlecol-new}
%%%%%%%%%%%%%%%%%%%%%%

\usepackage{natbib,stfloats}
\usepackage{mathrsfs,amsmath,upgreek}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{booktabs}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\def\newblock{\hskip .11em plus .33em minus .07em}

\newcommand{\be}{\begin{eqnarray}}
\newcommand{\ee}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}

\theoremstyle{TH}{
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corrolary}[lemma]{Corrolary}
\newtheorem{conjecture}[lemma]{Conjecture}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{claim}[lemma]{Claim}
\newtheorem{stheorem}[lemma]{Wrong Theorem}
%\newtheorem{algorithm}{Algorithm}
}

\theoremstyle{THrm}{
\newtheorem{definition}{Definition}
\newtheorem{question}{Question}
\newtheorem{remark}{Remark}
\newtheorem{scheme}{Scheme}
}

\theoremstyle{THhit}{
\newtheorem{case}{Case}[section]
}

\makeatletter

\def\Reals{\mathbb{R}}
\def\Ints{\mathbb{Z}}
\def\Nats{\mathbb{N}}

\def\theequation{\arabic{equation}}

\def\tc{\textcolor{red}}

\def\BottomCatch{%
\vskip -10pt
\thispagestyle{empty}%
\begin{table}[b]%
\NINE\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcr@{}}%
\\[-12pt]
Copyright \copyright\ 2012 Inderscience Enterprises Ltd. & &%
\end{tabular*}%
\vskip -30pt%
%%\vskip -35pt%
\end{table}%
%\def\tc{\textcolor}{red}
} \makeatother

%%%%%%%%%%%%%%%%%
\begin{document}%
%%%%%%%%%%%%%%%%%

\thispagestyle{plain}

\setcounter{page}{1}

\LRH{}

\RRH{}

\VOL{x}

\ISSUE{x}

\PUBYEAR{ }

%\BottomCatch

%\CLline

%\subtitle{}

\title{Optimised Cost Considering Huffman Code for Biological Data Compression}

%\authorA{xxxx}
%
%\affA{xxxx*\\
%*Corresponding author}
%
%\authorB{xxxx}
%\affB{xxxx}

\begin{abstract}
Classical Huffman code has widely been used to compress biological datasets. Though a considerable reduction of size of data is obtained by classical Huffman code, yet, more efficient encoding is possible by treating binary bits differently considering requirement of transmission time, energy consumption, and likewise. A number of techniques already modified Huffman code algorithm to  obtain optimal prefix-codes for unequal letter costs in order to reduce overall transmission cost (time). In this paper, we propose a new approach to improve compression performance of one such extension, cost considering approach (CCA), by applying genetic algorithm for optimal allocation of the codewords to the symbols. %The approach starts it operation by generating a set of codewords for different input symbols based on their frequencies by considering cost (length) of 0 and 1 as $\alpha$ and $\beta$,~$\alpha < \beta$. This step ensures optimal cost encoding, however, compression performance is not better than classical Huffman code. 
The idea of the proposed approach is to sacrifice some cost to minimise the total number of bits, hence, the genetic algorithm works by giving penalty on the cost. The performance of the approach is evaluated by applying it to compress some standard biological datasets. The experiments yield that the proposed approach improves the compression performance of the CCA considerably without increasing the cost significantly.
\end{abstract}

\KEYWORD{Data Compression, Huffman Code, Information Coding, Genetic Algorithm, Cost Considering Approach, Data Communication}

%\REF{to this paper should be made as follows: xxxx (xxxx) `xxxx',
%{\it xxxx}, Vol.~x, No.~x, pp.xxx--xxx.}

%\begin{bio}
%\tc{AUTHOR PLEASE SUPPLY CAREER HISTORY OF NO MORE THAN 100 WORDS
%FOR EACH AUTHOR.}
%\end{bio}

\maketitle

\section{Introduction}
\label{sec-introduction}
In the recent years, application of battery-powered portable devices, e.g. laptop computers and mobile phones has increased significantly. Proper representation of digital data and their transmission efficiency has become a primary concern for digital community because it affects the performance, reliability, and the cost of computation in both portable and non-portable devices. CMOS technologies were developed in order to reduce the power consumption both in data processing and transmission. In order to increase transmission speed and reduce transmission cost, parallel data transmission methods are widely used. However, parallel transmission is limited to short distance communications, e.g. locally connected devices, internal buses. Ruling out the possible availability of parallel transmission links over long distance, we are left with its serial alternative only. If we attempt to transfer big files, e.g. DNA sequences, over a serial transmission link then it would take a significant amount of time. However, we cannot overlook this problem because at present parallel processing is widely used to increase throughput and in parallel processing architecture, processing units are usually distributed in different physical locations and task sharing is a must in such architecture.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.8}
\caption{Example of binary logic level}
\label{table1}
%\centering
\begin{tabular}{ c  c  c }
\toprule
%\hline
 \bfseries Technology  & 0 & 1\\
\hline
\bfseries CMOS & $0~V$ to $\dfrac{V_{DD}}{2}$&$\dfrac{V_{DD}}{2}$ to $V_{DD}$\\
%\hline
\bfseries TTL & $0~V$ to $0.8~V$&$2~V$ to $V_{CC}$ \\
%\hline
%\hline
\bottomrule
\end{tabular}
\end{table}

@@

A recent deluge of interest in the development of new tools for biological data processing requires efficient methods for data compression. The main objective of data compression methods is minimising the number of bits in the data representation. \citet{bra09} proposed a new general data structure and data encoding approach for the efficient storage of genomic data. This method encodes only the differences between a genome sequence and a reference sequence. For encoding, the method uses different fixed-length encoding schemes such as Golomb \citep{Golomb96}, Elias codes \citep{Elias75} and variable length codes such as Huffman codes. There are other methods based on the same idea of encoding only the difference between reference sequence and the target one. One such approach~\citep{chr09} uses Huffman code for encoding difference between sequences to sent as an email attachment. The main limitation of this method is that it must send the reference sequence for at least once for each species, and usually this sequence is too big to be sent as an email attachment.
\citet{wan11} proposed a new scheme for referential compression of genomes based on the chromosome level. The algorithm searches for longest common subsequence between two sequences and then the differences between them are encoded using Huffman code.
All previous studies focus only on the differences and the relation between continuation of the sequence, and use existing encoding approaches to encode biological datasets without considering possible improvement of the encoding schemes.
%\subsection{Rationale of Unequal Bit Cost Considering Encoding Approaches}
\subsection{Huffman Codes}
In computer science and information theory, Huffman code is an entropy encoding algorithm used for lossless data compression. It takes into account the probabilities at which different symbols are likely to occur and results into fewer data bits on the average. 
%then assigns variable-length codes to different distinct symbols, and thus minimises the number of bits required to encode a message. 
For any given set of symbols and associated occurrence probabilities, there is an optimal encoding rule that minimises the number of bits needed to represent the source. Encoding symbols in predefined fixed length code does not attain an optimum performance, because every character consumes equal number of bits irrespective of their degree of contribution to the whole data. Huffman code tackles this by generating variable length codes, given a probability usage frequency for a set of symbols. It generates prefix-code to facilitate unambiguous retrieval of information. A scheme of prefix code assigns codes to letters in $\Sigma$ to form codeword $w_i$ such that none of them is a prefix to another. For example, the codes $\left\{ 1,01,001,0001\right\}$ and $\left\{ 000,001,011,111\right\}$ are prefix-free, whereas the code $\left\{ 1,01,100\right\}$ is not, because 1 is a prefix in 100.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.3}
\caption{Example of application of Huffman Code to compress ASCII characters}
\label{table2}
%\centering
\begin{tabular}{  c  c  c  c }
%\hline
\toprule
 Symbols & Frequency& ASCII Code & Huffman Code \\
\hline
\bfseries A & 50&01000001&00\\
%\hline
\bfseries B & 35&01000010&101\\
%\hline
\bfseries C & 42&01000011&110\\
%\hline
\bfseries D & 22&01000100&1001\\
%\hline
\bfseries E & 65&01000101&01\\
%\hline
\bfseries F & 25&01000110&1111\\
%\hline
\bfseries G & 9&01000111&1000\\
%\hline
\bfseries H & 23&01001000&1110\\
%\hline
\bottomrule
\end{tabular}
\end{table}

@@

\section{The Proposed Approach}
\label{sec3}
A genome is a stretch of DNA that encodes  a polypeptide (protein) which is a set of amino acids bound together in a specific order. Each genomic sequence consists of nucleotides bound together, which are interpreted by the cellular machinery in groups of three, called triplets~\citep{Harvey00}. This is the main reason to divide the whole sequence in a set of triplet and give a code to each triplet. As each DNA sequence contains a combination of four nucleobases\textemdash guanine (G), adenine (A), thymine (T), and cytosine (C), therefore it is possible to have $4^3=64$ triplets.  The first step in the optimised cost considering algorithm is cutting the genome sequence in triplets, then compute the frequency of each triplet in the whole sequence. This table of frequencies are used by the cost considering Huffman code to generate minimal cost code for each triplet (frequency). Finally, these codes with frequencies are used by the optimised cost considering algorithm to generate the optimal allocation with a given  penalty on cost. The whole process is shown in Fig.1.%\ref{Fig1}.

\begin{figure}[!thpb]
\caption{The proposed scheme}
\begin{center}
\includegraphics[scale=0.3]{Images/Drawing3_1.jpg}

\end{center}
\label{Fig1}
\end{figure}

@@

\begin{algorithm}[t]
\caption{Cost Considering Algorithm (CCA)}
\label{alg1}

@@

\subsection{Optimal Allocation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each triplet can be   modelled as an assignment problems, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $F=\left\{F_{1},F_{2}...F_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$, the objective is to assign to each frequency a code in order to minimise the total number of bits, while respecting the initially assigned total cost $S_{t}$ with a given penalty $\lambda\rightarrow [0,1]$. This penalty coefficient represents the allowed amount of cost that can be sacrificed to optimise the total number of bits.    
\\
The Objective Function is to:

@@

\subsubsection{Basic Genetic Algorithm}
Genetic Algorithm (GA) is a bio-inspired meta-heuristic algorithm developed by \citet{gen1}. GA is a stochastic optimisation algorithm imitate the natural evolution process of genomes. GA started by generating an initial population of random feasible solutions. The optimisation process of GA takes the initial population and generates a new population based on it.  The process can be described as follow:\\ 
Firstly, select two or more solutions from the current population by using one of the well-known selection techniques \citep{bli95}. These selected solutions will be considered as  parents. The second operation of the genetic algorithm is the crossover, which takes these parents as inputs to generate other new solutions considered as children. The third operation of the genetic algorithm is the mutation which ensures a good diversification in the search process. The new solutions can be mutated according to a given mutation probability. The mutation operation changes the value of one or more positions in the solution. The quality of each solution is verified by the fitness function which controls the evolution of the GA population by deletion of the worst solutions and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is reached which can be the number of generation or if the population is stabilised.

\subsubsection{Optimised cost considering algorithm}
The main objective of the optimised cost considering algorithm (OCCA) is to apply GA to find an optimal allocation of the codewords to each triplets to reduce the total number of bits. In GA, the encoding of the solution into a chromosome is one of the most important issues in obtaining good optimisation results. The OCCA uses two fixed length arrays of size ``64" which is the number of combination for all nucleotides. The first array contains the frequencies of each triplets and the second array contains the cost of the codewords assigned to each triplets. Our genetic algorithm utilises the two arrays and uses the index of each entries in the allocation process.

@@

@@

The evolution of the population is the key of the genetic optimisation algorithm. During each generation the process start with the selection of a part (set of solutions) of the population to bread a new generation. In the literature many selection methods have been proposed to guide the population evolution \citep{bli95}. The different methods for selecting a part (set of solutions) of the population vary from a random selection method to heuristic based selection method. We have chosen to select randomly the part of the population to be processed as the heuristic methods are very time consuming (step 3).
After that the operations of genetic algorithm are applied on the initial population to generate a new generation of the population (see Fig.2). Firstly the crossover operation is applied to these two selected solutions (considered as parents) to generate two new solutions (considered as children)(step 4). In the literature many crossover techniques have been used in genetic algorithm \citep{osa14}, such as one-point crossover which divides the chromosomes into two fragments and recombines the second fragments with the other chromosome's second fragment. The two-point crossover divides the chromosomes into three fragments and recombines the middle fragment with  the middle fragment of the other chromosome. There are many other crossover techniques to allow a good convergence of the algorithm. In our case, we have used the two point crossover with two parameters, the first parameter $p_{c}$ is a random value in [0,63], which represents the first cut point and the second parameter $p_{m}$ is also a random value in [0,63], which represents the number of positions to be crossed. These two random parameters are to ensure a good diversification on the whole search space (step 4).

\begin{figure}[h]
\caption{Operations of genetic algorithm}
\begin{center}
\includegraphics[scale=0.4]{Images/Drawing2.jpg}
\end{center}
\label{Fig3}
\end{figure}

\begin{figure}[h]
\caption{Population update for genetic algorithm}
\begin{center}
\includegraphics[scale=0.45]{Images/Drawing_1.jpg}

\end{center}
\label{Fig2}
\end{figure}

After the crossover operation, the newly generated children may contain conflict, for example, a single code is allocated to two different frequencies in the solution. To overcome such a conflict, a regulation operation is performed to refine the solution to ensure the correctness of the solution (see Fig.2).
 Secondly these two new solutions are mutated according to a predefined probability $\gamma$, the best value of the mutation rate is  problem specific (step 5). In our case, the value of $\gamma$ is fixed to $0.2$ to explore a few positions in the solution. The mutation operator is used to maintain genetic diversity from one generation of the population of genetic algorithm to the next. In our case, we have used the mutation as a random swap mutation operator (see Fig.3).  Each newly generated solution must satisfy the cost constraint. The next step is to add these two new solutions (children) to the current population (step 7) (see Fig. 3). Finally, the new population are ranked by fitness (step 8), and the worst solutions are deleted until the initial size of the population is obtained  (step 9). The whole processes are repeated until the maximum number of operation is performed (step 10).

@@

\begin{figure}[h]
\caption{Convergence of OCCA for \textit{Genome 2}}
\begin{center}
\includegraphics[scale=0.4]{Images/drawing4_1.jpg}
\end{center}
\label{Fig5}
\end{figure}

\begin{table*}[!thpb]
\renewcommand{\arraystretch}{1.3}
\small
\label{table3}
\caption{Datasets description}
%\begin{center}
%\centering
\begin{tabular}{ c  c  c  c }
%\hline
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MB)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &  6.66 & CP009496  \\
%\hline

Genome 2& Amycolatopsis benzoatilytica & 8.30 & NZ\_KB912942 \\
%\hline

@@

%\hline

Genome 4 & \begin{tabular}{@{}c@{}}Streptomyces bottropensis\\ATCC 25435\end{tabular}& 8.54 &
NZ\_KB911581\\
%\hline

Genome 5 & \begin{tabular}{@{}c@{}}Mycobacterium smegmatis\\str. MC2 155\end{tabular}& 6.66 &
CP009494            \\
%\hline

Genome 6 & \begin{tabular}{@{}c@{}}Mycobacterium smegmatis\\MKD8\end{tabular}& 6.76 &
NZ\_KI421511            \\
%\hline

Genome 7& Bradyrhizobium WSM471&  7.42 &NZ\_CM001442\\%\hline

Genome 8 & \begin{tabular}{@{}c@{}}Amycolatopsis thermoflava\\N1165\end{tabular}& 8.27 &
NZ\_CM001442 \\
%\hline

Genome 9 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\Bt407\end{tabular}& 5.74 &
NZ\_CM000747 \\
%\hline

Genome 10 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar thuringiensis\end{tabular}& 6.03 & NZ\_CM000748 \\
%\hline

Genome 11 & \begin{tabular}{@{}c@{}}Pseudomonas aeruginosa\\9BR\end{tabular}& 6.48 & NZ\_AFXI01000001 \\
%\hline

Genome 12 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar berliner ATCC\end{tabular}& 5.97 & NZ\_CM000753 \\
%\hline

Genome 13 & \begin{tabular}{@{}c@{}}Bacillus thuringiensis\\serovar pakistani\end{tabular}& 5.75 & NZ\_CM000750 \\
%\hline

Genome 14 & \begin{tabular}{@{}c@{}}Pseudomonas aeruginosa\\LES400\end{tabular}& 6.28 & CP006982 \\
%\hline

Genome 15 & \begin{tabular}{@{}c@{}}Mus musculus\\chromosome 1\end{tabular}& 25.58 & GL456087 \\
%\hline

Genome 16 & \begin{tabular}{@{}c@{}}Danio rerio\\chromosome 1\end{tabular}& 56.14 & CM002885 \\
%\hline

Genome 17 & \begin{tabular}{@{}c@{}}Homo sapiens\\chromosome 18\end{tabular}& 76.64 & CM000680  \\
%\hline

Genome 18 & \begin{tabular}{@{}c@{}}Homo sapiens\\chromosome 22\end{tabular}& 99.94 & CM000684    \\
%\hline

\bottomrule
\end{tabular}
%\end{center}
\end{table*}

\begin{table*}[!thpb]
\renewcommand{\arraystretch}{1.2}
\small
\label{table4}
\caption{Comparison of performance among classical Huffman code, CCA, and OCCA without penalty}
%\begin{center}
